Model Checking Differentially Private Properties

### 2018-06-08 preliminaries.tex
polish

\subsection{Differential Privacy}
We denote the data universe by $\mathcal{X}$; $x \in \mathcal{X}^n$ is
a dataset with $n$ rows from the data universe. Two datasets $x$ and
$x'$ are \emph{neighbors} (denoted by $d(x, x') \leq 1$) if they are
identical except at most one row. A \emph{query} $f$ is a function
from $\mathcal{X}^n$ to its range $R \subseteq \bbfZ$. The \emph{sensitivity} of the
query $f$ (written $\Delta (f)$) is $\max_{d(x, x') \leq 1} | f (x) -
f (x') |$. For instance, a \emph{counting} query counts the number
of rows with certain attributes (say, female). The sensitivity of a
counting query is $1$ since any neighbor can change the count by at
most one. We only consider queries with finite  ranges for simplicity.
A \emph{data analysis mechanism} (or
\emph{mechanism} for brevity) $M_f$ for a query $f$
is a randomized algorithm with inputs in $\mathcal{X}^n$ and outputs
in $\tilde{R}$.
A mechanism may not have the same output range as its query, that is,
$\tilde{R} \neq R$ in general.
A mechanism $M_f$ for $f$ is \emph{oblivious} if
$\Pr[M_f(x) = \tilde{r}] = \Pr[M_f(x') = \tilde{r}]$ for every
$\tilde{r} \in \tilde{R}$ when $f (x) = f (x')$. In words, outputs of an
oblivious mechanism depend on the query result $f (x)$.
The order of rows in a dataset, for instance, is
irrelevant to oblivious mechanisms. Let $x, x'$
be datasets and $\tilde{r} \in \tilde{R}$. We say the probability of
the mechanism $M_f$ outputting $\tilde{r}$ on $x$ is
\emph{$(\epsilon, \delta)$-close} to those on $x'$ if
\[
\Pr[M_f (x) = \tilde{r}] \leq e^{\epsilon} \Pr[M_f (x') =
\tilde{r}] + \delta.
%\footnote{The standard definition requires
%$\Pr[M_f (x) = \tilde{r}] \leq e^\epsilon \Pr[M_f (x') =
%\tilde{r}] + \delta$~\cite{DMNS:06:CNSPD,D:06:DP}.
%Since $\epsilon = e^{\ln \epsilon}$, our $(\epsilon, \delta)$-closedness is
%equivalent to the standard $(\ln \epsilon,
%\delta)$-closedness~\cite{GRS:09:UUPM,GRS:12:UUPM}.}
\]
A mechanism $M_f$ is \emph{$(\epsilon, \delta)$-differentially
  private} % $(\epsilon, \delta \approx 0)$
if for every $x, x' \in \mathcal{X}^n$ with $d(x, x') \leq 1$
and $\tilde{r} \in \tilde{R}$, the probability of
$M_f$ outputting $\tilde{r}$ on $x$ is $(\epsilon, \delta)$-close to
those on $x'$.

The non-negative parameters $\epsilon$ and $\delta$ quantify
mechanism behaviors probabilistically;
the smaller they are, the behaviors are more similar.
Informally, a differentially private mechanism has probabilistically
similar behaviors on neighboring datasets. It will have similar
output distributions when any row is replaced by another in a given
dataset. 
%Consider, for instance, a database contains the row
%corresponding to an individual. A differentially private mechanism
%will have a similar output distribution when the individual is not in
%the database. Such mechanisms behave similarly regardless of the
%presence of any individual. 
Individual privacy is thus preserved by differentially private mechanisms.

@@

Let $M = (S, \Act, \wp, L)$ be an MDP. A (history-dependent)
\emph{scheduler} for $M$ is a function $\scheduler{S} : S^+
\rightarrow \Act$. A \emph{query scheduler} for $M$ is a function
$\scheduler{Q} : S^+ \rightarrow \Act$ such that
$\scheduler{Q}(\sigma) = \scheduler{Q}(\sigma')$ for any $\sigma,
\sigma' \in S^+$ of the same length.
Intuitively, its decision depends only on the length of the history.
A path $\pi =
\pi_0 \alpha_1 \pi_1 \cdots$ $\pi_n$ $\alpha_{n+1} \cdots$ is an
\emph{$\scheduler{S}$-path} if $\alpha_{i+1} = \scheduler{S}(\pi_0
\pi_1 \cdots \pi_i)$ for all $i \geq 0$.
Note that an MDP with a
scheduler $\scheduler{S}$ induces a Markov chain $M_{\scheduler{S}} =
(S^+, \wp_{\scheduler{S}}, L')$ where $L' (\sigma s) = L (s)$,
$\wp_{\scheduler{S}} (\sigma s, \sigma s t) = \wp (s,
\scheduler{S}(\sigma s), t)$ for $\sigma \in S^*$ and $s, t \in S$.
\hide{
Subsequently, we write
$\Pr_{\scheduler{S}} (\Pi)$ for the \emph{probability measure} of the
measurable path set $\Pi$ on $M$ given the scheduler $\scheduler{S}$.
}

>>>>

\subsection{Differential Privacy}
We denote the data universe by $\mathcal{X}$; $x \in \mathcal{X}^n$ is
a \emph{dataset} with $n$ \emph{rows} from the data universe. Two datasets $x$ and
$x'$ are \emph{neighbors} (denoted by $d(x, x') \leq 1$) if they are
identical except at most one row. A \emph{query} $f$ is a function
from $\mathcal{X}^n$ to its range $R \subseteq \bbfZ$. The \emph{sensitivity} of the
query $f$ (written $\Delta (f)$) is $\max_{d(x, x') \leq 1} | f (x) -
f (x') |$. For instance, a \emph{counting} query counts the number
of rows with certain attributes (say, female). The sensitivity of a
counting query is $1$ since any neighbor can change the count by at
most one. We only consider queries with finite  ranges for simplicity.
A \emph{data analysis mechanism} (or
\emph{mechanism} for brevity) $M_f$ for a query $f$
is a randomized algorithm with inputs in $\mathcal{X}^n$ and outputs
in $\tilde{R}$.
A mechanism may not have the same output range as its query, that is,
$\tilde{R} \neq R$ in general.
A mechanism $M_f$ for $f$ is \emph{oblivious} if
$\Pr[M_f(x) = \tilde{r}] = \Pr[M_f(x') = \tilde{r}]$ for every
$\tilde{r} \in \tilde{R}$ when $f (x) = f (x')$. In words, outputs of an
oblivious mechanism depend on the query result $f (x)$.
The order of rows, for instance, is
irrelevant to oblivious mechanisms. Let $x, x'$
be datasets and $\tilde{r} \in \tilde{R}$. The probability of
the mechanism $M_f$ outputting $\tilde{r}$ on $x$ is
\emph{$(\epsilon, \delta)$-close} to those on $x'$ if
\[
\Pr[M_f (x) = \tilde{r}] \leq e^{\epsilon} \Pr[M_f (x') =
\tilde{r}] + \delta.
%\footnote{The standard definition requires
%$\Pr[M_f (x) = \tilde{r}] \leq e^\epsilon \Pr[M_f (x') =
%\tilde{r}] + \delta$~\cite{DMNS:06:CNSPD,D:06:DP}.
%Since $\epsilon = e^{\ln \epsilon}$, our $(\epsilon, \delta)$-closedness is
%equivalent to the standard $(\ln \epsilon,
%\delta)$-closedness~\cite{GRS:09:UUPM,GRS:12:UUPM}.}
\]
A mechanism $M_f$ is \emph{$(\epsilon, \delta)$-differentially
  private} % $(\epsilon, \delta \approx 0)$
if for every $x, x' \in \mathcal{X}^n$ with $d(x, x') \leq 1$
and $\tilde{r} \in \tilde{R}$, the probability of
$M_f$ outputting $\tilde{r}$ on $x$ is $(\epsilon, \delta)$-close to
those on $x'$.

The non-negative parameters $\epsilon$ and $\delta$ quantify
mechanism behaviors probabilistically;
the smaller they are, the behaviors are more similar.
Informally, a differentially private mechanism has probabilistically
similar behaviors on neighbors. It will have similar
output distributions when any row is replaced by another in a given
dataset. 
%Consider, for instance, a database contains the row
%corresponding to an individual. A differentially private mechanism
%will have a similar output distribution when the individual is not in
%the database. Such mechanisms behave similarly regardless of the
%presence of any individual. 
Since the output distribution does not change significantly with the
absence of any row in a dataset,
individual privacy is thus preserved by differentially private mechanisms.

@@

Let $M = (S, \Act, \wp, L)$ be an MDP. A (history-dependent)
\emph{scheduler} for $M$ is a function $\scheduler{S} : S^+
\rightarrow \Act$. A \emph{query scheduler} for $M$ is a function
$\scheduler{Q} : S^+ \rightarrow \Act$ such that
$\scheduler{Q}(\sigma) = \scheduler{Q}(\sigma')$ for any $\sigma,
\sigma' \in S^+$ of the same length.
Intuitively, decisions of a query scheduler depend only on the length of the history.
A path $\pi =
\pi_0 \alpha_1 \pi_1 \cdots$ $\pi_n$ $\alpha_{n+1} \cdots$ is an
\emph{$\scheduler{S}$-path} if $\alpha_{i+1} = \scheduler{S}(\pi_0
\pi_1 \cdots \pi_i)$ for all $i \geq 0$.
Note that an MDP with a
scheduler $\scheduler{S}$ induces a Markov chain $M_{\scheduler{S}} =
(S^+, \wp_{\scheduler{S}}, L')$ where $L' (\sigma s) = L (s)$,
$\wp_{\scheduler{S}} (\sigma s, \sigma s t) = \wp (s,
\scheduler{S}(\sigma s), t)$ for $\sigma \in S^*$ and $s, t \in S$.
\hide{
Subsequently, we write
$\Pr_{\scheduler{S}} (\Pi)$ for the \emph{probability measure} of the
measurable path set $\Pi$ on $M$ given the scheduler $\scheduler{S}$.
}

### 2018-06-08 semantics-mdp.tex
polish

\noindent
\paragraph{Justification of query schedulers.}
We use query schedulers in the semantics for the differentially
private operator. A definition with history-dependent schedulers
might be
\begin{eqnarray*}
  M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi
  & \textmd{ if } &
  \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
  scheduler } \scheduler{S},
  \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
  e^{\epsilon} \cdot\\
  && \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta \textmd{ and }
  \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
  e^{\epsilon} \cdot \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
  + \delta
\end{eqnarray*}
A state satisfies $\dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi$ if
no history-dependent scheduler can differentiate the probabilities
of having paths satisfying
$\phi$ from neighbors. Recall that a history-dependent scheduler takes
actions by previous states. Such a definition would allow
schedulers to take different actions from different states. 
Two neighbors could hence be differentiated by different action sequences.
The specification might be too strong for our purposes.
A query scheduler $\scheduler{Q} : S^+ \rightarrow \Act$,
on the other hand, corresponds to a query sequence. A state satisfies
$\dpriv{\epsilon}{\delta} \phi$ if no query sequence can differentiate
the probabilities of having paths satisfying $\phi$ from neighbors.
Recall query schedulers only depend on lengths of histories. Two
neighbors cannot be distinguished by the same action sequence of any
length if they satisfy a differentially private subformula.
Our semantics agrees with the informal interpretation of differential
privacy for such systems. We therefore consider only query schedulers
in our definition.

\subsection{Model Checking}
Given an MDP $M =
(S, \Act, \wp, L)$, a neighborhood relation $\neighbor{S}$, $s \in
S$, and a path formula $\phi$, consider the problem of deciding $M,
\neighbor{S}, s \models 
\dpriv{\epsilon}{\delta} \phi$ . Recall the semantics of $\dpriv{\epsilon}{\delta}
\phi$. Given $s, t$ with $s \neighbor{S} t$ and a path formula
$\phi$, we need to decide whether
$\myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} \leq
e^{\epsilon} \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} + \delta$
for every query scheduler $\scheduler{Q}$.
When $\phi$ is $\gX B$ with $B\subseteq S$, only the first action in
the query sequence needs to be considered. This can also be easily generalized to nested next operators:  one needs only to enumerate
all actions query sequences of a fixed length.
% obtained by the number of nested $\X$ operators.
The problem however is undecidable in general.

@@

We model the above threshold algorithm (Algorithm~\ref{algorithm:online-model})
and apply
Proposition~\ref{proposition:sufficient-condition-mdp-model-checking} 
check whether the mechanism is differentially private using the
classical $\PCTL$ model checking algorithm for MDPs. 
Since concrete values of the parameters $\epsilon$ and $\delta$ are
computed, tighter bounds for specific neighbors can be obtained.
For instance, for the state $t_3r_5$ and its neighbor
$\underline{t}_3\underline{r}_4$, we verify the property $\bigwedge_{k
  \in \bbfZ^{\geq 0}} \dpriv{0}{0.17} ((\X^k\bot) \top)$ is
satisfied. Note the reachability probability goes to $0$ as $k$ goes
to infinity. By repeating the computation, we verify that 
the property $\bigwedge_{k \in \bbfZ^{\geq 0}} \dpriv{1}{0.74}
((\X^k\bot) \top)$ is satisfied for all neighbors. 
Subsequently, the above threshold mechanism in
Algorithm~\ref{algorithm:online-model} is $(1, 0.74)$-differentially
private. Compared to the parameters for the neighbors $t_3r_5$ and
$\underline{t}_3\underline{r}_4$, the parameter $\delta$ appears to be
significantly large. It means that there are two neighbors with
drastically different output distributions from our mechanism.
Moreover, recall that
Proposition~\ref{proposition:sufficient-condition-mdp-model-checking}
is a sufficient condition. It only gives an upper bound of privacy
parameters. Automatically computing tighter bounds may be possible.

>>>>

\noindent
\paragraph{Justification of query schedulers.}
We use query schedulers in the semantics for the differentially
private operator. A definition with history-dependent schedulers
might be
\begin{eqnarray*}
  M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi
  & \textmd{ if } &
  \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
  scheduler } \scheduler{S},
  \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
  e^{\epsilon} \cdot\\
  && \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta \textmd{ and }
  \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
  e^{\epsilon} \cdot \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
  + \delta.
\end{eqnarray*}
A state satisfies $\dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi$ if
no history-dependent scheduler can differentiate the probabilities
of having paths satisfying
$\phi$ from neighbors. Recall that a history-dependent scheduler chooses
actions according to previous states. Such a definition would allow
schedulers to take different actions from different states. 
Two neighbors could hence be differentiated by different action sequences.
The specification might be too strong for our purposes.
A query scheduler $\scheduler{Q} : S^+ \rightarrow \Act$,
on the other hand, corresponds to a query sequence. A state satisfies
$\dpriv{\epsilon}{\delta} \phi$ if no query sequence can differentiate
the probabilities of having paths satisfying $\phi$ from neighbors.
Recall query schedulers only depend on lengths of histories. Two
neighbors cannot be distinguished by the same action sequence of any
length if they satisfy a differentially private subformula.
Our semantics agrees with the informal interpretation of differential
privacy for such systems. We therefore consider only query schedulers
in our definition.

\subsection{Model Checking}
Given an MDP $M =
(S, \Act, \wp, L)$, a neighborhood relation $\neighbor{S}$, $s \in
S$, and a path formula $\phi$, consider the problem of checking $M,
\neighbor{S}, s \models 
\dpriv{\epsilon}{\delta} \phi$ . Recall the semantics of $\dpriv{\epsilon}{\delta}
\phi$. Given $s, t$ with $s \neighbor{S} t$ and a path formula
$\phi$, we need to decide whether
$\myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} \leq
e^{\epsilon} \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} + \delta$
for every query scheduler $\scheduler{Q}$.
When $\phi$ is $\gX B$ with $B\subseteq S$, only the first action in
the query sequence needs to be considered. This can also be easily generalized to nested next operators:  one needs only to enumerate
all actions query sequences of a fixed length.
% obtained by the number of nested $\X$ operators.
The problem however is undecidable in general.

@@

We model the above threshold algorithm (Algorithm~\ref{algorithm:online-model})
and apply
Proposition~\ref{proposition:sufficient-condition-mdp-model-checking} 
to check whether the mechanism is differentially private using the
classical $\PCTL$ model checking algorithm for MDPs. 
Since concrete values of the parameters $\epsilon$ and $\delta$ are
computed, tighter bounds for specific neighbors can be obtained.
For instance, for the state $t_3r_5$ and its neighbor
$\underline{t}_3\underline{r}_4$, we verify the property $\bigwedge_{k
  \in \bbfZ^{\geq 0}} \dpriv{0}{0.17} ((\X^k\bot) \top)$ is
satisfied. Note the reachability probability goes to $0$ as $k$ goes
to infinity. By repeating the computation, we verify that 
the property $\bigwedge_{k \in \bbfZ^{\geq 0}} \dpriv{1}{0.74}
((\X^k\bot) \top)$ is satisfied for all neighbors. 
Subsequently, the above threshold mechanism in
Algorithm~\ref{algorithm:online-model} is $(1, 0.74)$-differentially
private. Compared to the parameters for the neighbors $t_3r_5$ and
$\underline{t}_3\underline{r}_4$, the parameter $\delta$ appears to be
significantly large. It means that there are two neighbors with
drastically different output distributions from our mechanism.
Moreover, recall that
Proposition~\ref{proposition:sufficient-condition-mdp-model-checking}
is a sufficient condition. It only gives an upper bound of privacy
parameters. Tighter bounds may be computed by more sophisticated
sufficient conditions.

### 2018-06-08 semantics.tex
polish

Given a Markov chain $K = (S, \wp, L)$, a
\emph{neighborhood relation} $\neighbor{S} \subseteq S \times S$
is a reflexive and symmetric relation on $S$. We will write $s
\neighbor{S} t$ when $(s, t) \in \neighbor{S}$. If $s \neighbor{S} t$,
we say $s$ and $t$ are \emph{neighbors} or $t$ is a \emph{neighbor} of
$s$. Define the satisfaction relation $K, \neighbor{S}, s
\models \Phi$ as follows.
\begin{eqnarray*}
%  K, \neighbor{S}, s \models \top\\
  K, \neighbor{S}, s \models p
  & \textmd{ if } &
  p \in L(s)\\
  K, \neighbor{S}, s \models \neg \Phi
  & \textmd{ if } &
  K, \neighbor{S}, s \not\models \Phi\\
  K, \neighbor{S}, s \models \Phi_0 \wedge \Phi_1
  & \textmd{ if } &
  K, \neighbor{S}, s \models \Phi_0 \textmd{ and }
  K, \neighbor{S}, s \models \Phi_1\\
  K, \neighbor{S}, s \models \PJ{J} \phi
  & \textmd{ if } &
  \Pr [\{ \pi : K, \neighbor{S}, \pi \models \phi \textmd{ with }
                    \pi_0 = s\}] \in J \\
  K, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi
  & \textmd{ if } &
  \textmd{for every } t \textmd{ with }  s \neighbor{S} t,
      p(s) \leq e^{\epsilon} p(t) + \delta \textmd{ and }
      p(t) \leq e^{\epsilon} p(s)  +  \\
  & &  \delta \textmd{ where } p(x) = \Pr [\{
      \pi : K, \neighbor{S}, \pi \models \phi \textmd{ with }
      \pi_0 = x \}]
\end{eqnarray*}

Moreover, the relation $K, \neighbor{S}, \pi \models \phi$ is defined as
for the standard linear temporal logic formulae~\cite{ManneP92}. We only recall the semantics for $\X$ and $\until$:
\begin{eqnarray*}
  K, \neighbor{S}, \pi \models \X \phi
  & \textmd{ if } &
  K, \neighbor{S}, \pi[1] \models \phi\\
  K, \neighbor{S}, \pi \models \phi \until \psi
  & \textmd{ if } &
  \textmd{there is a } j \geq 0 \textmd{ such that }
  K, \neighbor{S}, \pi[j] \models \psi \textmd{ and } \\
  & & K, \neighbor{S}, \pi[k] \models \phi
      \textmd{ for every } 0 \leq k < j
\hide{
  K, \neighbor{S}, \pi \models \Phi \buntil{n} \Psi
  & \textmd{ if } &
  \textmd{there is a } 0 \leq j \leq n \textmd{ such that }
  K, \neighbor{S}, \pi[j] \models \Psi \textmd{ and } \\
  & & K, \neighbor{S}, \pi[k] \models \Phi
      \textmd{ for every } 0 \leq k < j
}
\end{eqnarray*}

For any Markov chain $K$, neighborhood
relation $N$ on $S$, $s \in S$, and a path formula $\phi$,
define
\[
\myPr{s}{K}{N}{\phi} =
\Pr [ \{ \pi : K, N, \pi \models \phi \textmd{ with } \pi_0 = s \} ].
\]
That is, $\myPr{s}{K}{N}{\phi}$ denotes the probability of paths
satisfying $\phi$ from $s$ on $K$ with $N$. The semantics of the
differentially private operator hence becomes
\begin{eqnarray*}
  K, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi
  & \textmd{ if } &
  \textmd{for every } t \textmd{ with }  s \neighbor{S} t,
      \myPr{s}{K}{\neighbor{S}}{\phi} \leq e^{\epsilon}
      \myPr{t}{K}{\neighbor{S}}{\phi} + \delta
      \textmd{ and }\\
  & &  \myPr{t}{K}{\neighbor{S}}{\phi} \leq e^{\epsilon}
       \myPr{s}{K}{\neighbor{S}}{\phi} + \delta
\end{eqnarray*}

@@

\begin{proposition}
  Let $K = (S, \wp, L)$ be a Markov chain and $\neighbor{S}$ a
  neighborhood relation on $S$.
  $\{ s : K, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi \}$ is
  computable within time polynomial in $|S|$.
\end{proposition}

>>>>

Given a Markov chain $K = (S, \wp, L)$, a
\emph{neighborhood relation} $\neighbor{S} \subseteq S \times S$
is a reflexive and symmetric relation on $S$. We will write $s
\neighbor{S} t$ when $(s, t) \in \neighbor{S}$. If $s \neighbor{S} t$,
we say $s$ and $t$ are \emph{neighbors} or $t$ is a \emph{neighbor} of
$s$. 
For any Markov chain $K$, neighborhood
relation $N$ on $S$, $s \in S$, and a path formula $\phi$,
define
\[
\myPr{s}{K}{N}{\phi} =
\Pr [ \{ \pi : K, N, \pi \models \phi \textmd{ with } \pi_0 = s \} ].
\]
That is, $\myPr{s}{K}{N}{\phi}$ denotes the probability of paths
satisfying $\phi$ from $s$ on $K$ with $N$. 
Define the satisfaction relation $K, \neighbor{S}, s
\models \Phi$ as follows.
\begin{eqnarray*}
%  K, \neighbor{S}, s \models \top\\
  K, \neighbor{S}, s \models p
  & \textmd{ if } &
  p \in L(s)\\
  K, \neighbor{S}, s \models \neg \Phi
  & \textmd{ if } &
  K, \neighbor{S}, s \not\models \Phi\\
  K, \neighbor{S}, s \models \Phi_0 \wedge \Phi_1
  & \textmd{ if } &
  K, \neighbor{S}, s \models \Phi_0 \textmd{ and }
  K, \neighbor{S}, s \models \Phi_1\\
  K, \neighbor{S}, s \models \PJ{J} \phi
  & \textmd{ if } &
                    \myPr{s}{K}{\neighbor{S}}{\phi} \in J\\
  K, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi
  & \textmd{ if } &
  \textmd{for every } t \textmd{ with }  s \neighbor{S} t,
      \myPr{s}{K}{\neighbor{S}}{\phi} \leq e^{\epsilon}
      \myPr{t}{K}{\neighbor{S}}{\phi} + \delta
      \textmd{ and }\\
  & &  \myPr{t}{K}{\neighbor{S}}{\phi} \leq e^{\epsilon}
       \myPr{s}{K}{\neighbor{S}}{\phi} + \delta
\end{eqnarray*}

Moreover, the relation $K, \neighbor{S}, \pi \models \phi$ is defined as
in the standard linear temporal logic formulae~\cite{ManneP92}. We only recall the semantics for the temporal operators $\X$ and $\until$:
\begin{eqnarray*}
  K, \neighbor{S}, \pi \models \X \phi
  & \textmd{ if } &
  K, \neighbor{S}, \pi[1] \models \phi\\
  K, \neighbor{S}, \pi \models \phi \until \psi
  & \textmd{ if } &
  \textmd{there is a } j \geq 0 \textmd{ such that }
  K, \neighbor{S}, \pi[j] \models \psi \textmd{ and } \\
  & & K, \neighbor{S}, \pi[k] \models \phi
      \textmd{ for every } 0 \leq k < j
\hide{
  K, \neighbor{S}, \pi \models \Phi \buntil{n} \Psi
  & \textmd{ if } &
  \textmd{there is a } 0 \leq j \leq n \textmd{ such that }
  K, \neighbor{S}, \pi[j] \models \Psi \textmd{ and } \\
  & & K, \neighbor{S}, \pi[k] \models \Phi
      \textmd{ for every } 0 \leq k < j
}
\end{eqnarray*}

@@

\begin{proposition}
  Let $K = (S, \wp, L)$ be a Markov chain, $\neighbor{S}$ a
  neighborhood relation on $S$, and $\phi$ a $\dpCTL$ path formula.
  $\{ s : K, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi \}$ is
  computable within time polynomial in $|S|$ and $|\phi|$.
\end{proposition}
